---
title: "01_Spatial_coverage"
output: html_document
---

## 1. Lib
```{r setup, include=FALSE}
library(tidyverse)
```

## 2. Read data
```{r}
dat_sed <- readxl::read_excel("Input_data/sediment_summary.xlsx", sheet = 2)
dat_bio <- readxl::read_excel("Input_data/biota_summary_Rob_2018-11-10.xlsx", sheet = 2)

# xtabs(~shape + colour, dat_sed)
# xtabs(~region, dat_bio) %>% names() %>% dput()
# xtabs(~detGroup, dat_bio) %>% names() %>% dput()
# xtabs(~detGroup, dat_sed) %>% names() %>% dput()

reg_ordered <- c("Barents Sea", "Greenland-Scotland ridge", "Norwegian Sea", 
  "Northern North Sea", "Southern North Sea", "Norwegian Trench", "Skagerrak and Kattegat", "Channel", 
  "Irish and Scottish West Coast", "Irish Sea", "Celtic Sea", 
  "Gulf of Cadiz", "Iberian Sea", "Northern Bay of Biscay")


detgroup_ordered <- c("Metals", "Chlorobiphenyls", "PAH (parent)", "Organo-bromines", 
                      "Organo-metals", "Imposex", "PAH (alkylated)", "Biological effects", 
                      "Dioxins", "Perfluoronated compounds", "Pesticides")

dat_sed$region <- factor(dat_sed$region, levels = reg_ordered)
dat_sed$detGroup <- factor(dat_sed$detGroup, levels = detgroup_ordered)

dat_bio$region <- factor(dat_bio$region, levels = reg_ordered)
dat_bio$detGroup <- factor(dat_bio$detGroup, levels = detgroup_ordered)


```

## 3. Sediment

### a. Tabulate number of stations including informal assessment    
```{r}
df1 <- dat_sed %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  count(OSPARregion, region, detGroup, determinand) 
# df1 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df2 <- df1 %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(n))
# df2 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df_sed_all <- df2 %>%
  spread(detGroup, N_stations_best)

df_sed_all
```

### b. Stations with status, spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_sed_status <- dat_sed %>%
  filter(!detGroup %in% "PAH (alkylated)" & !shape %in% "small_open_circle") %>%
  mutate(lat_r = round(latitude, 1), lon_r = 2*round(latitude/2, 1)) %>%
  mutate(Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, Location, detGroup, determinand) %>%
  count(OSPARregion, region, detGroup, determinand) %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(nn)) %>%
  spread(detGroup, N_stations_best)
```

### c. Stations with trend and status, spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_sed_trendstatus <- dat_sed %>%
  filter(!detGroup %in% "PAH (alkylated)" & !shape %in% c("small_open_circle", "small_filled_circle")) %>%
  mutate(lat_r = round(latitude, 1), lon_r = 2*round(latitude/2, 1)) %>%
  mutate(Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, Location, detGroup, determinand) %>%
  count(OSPARregion, region, detGroup, determinand) %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(nn)) %>%
  spread(detGroup, N_stations_best)
```



## 4. Biota  

### a. Tabulate number of stations including informal assessment    
```{r}
df1 <- dat_bio %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  count(OSPARregion, region, detGroup, determinand) 
# df1 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df2 <- df1 %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(n))
# df2 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df_bio_all <- df2 %>%
  spread(detGroup, N_stations_best)

df_bio_all
```

### b. Stations with status, spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_bio_status <- dat_bio %>%
  filter(!detGroup %in% "PAH (alkylated)" & !shape %in% "small_open_circle") %>%
  mutate(lat_r = round(latitude, 1), lon_r = 2*round(latitude/2, 1)) %>%
  mutate(Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, Location, detGroup, determinand) %>%
  count(OSPARregion, region, detGroup, determinand) %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(nn)) %>%
  spread(detGroup, N_stations_best)
```

### c. Stations with trend and status, spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_bio_trendstatus <- dat_bio %>%
  filter(!detGroup %in% "PAH (alkylated)" & !shape %in% c("small_open_circle", "small_filled_circle")) %>%
  mutate(lat_r = round(latitude, 1), lon_r = 2*round(latitude/2, 1)) %>%
  mutate(Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, Location, detGroup, determinand) %>%
  count(OSPARregion, region, detGroup, determinand) %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(nn)) %>%
  spread(detGroup, N_stations_best)
```

## 5. Save as excel

```{r}
df_list <- list(
  df_sed_all, df_sed_status, df_sed_trendstatus,
  df_bio_all, df_bio_status, df_bio_trendstatus
)

names(df_list) <- c(
  "sed_all", "sed_status", "sed_trendstatus",
  "bio_all", "bio_status", "bio_trendstatus"
)

openxlsx::write.xlsx(df_list, "Produced_data/01_Spatial_coverage.xlsx")

?extract

```

