--
title: "01_Spatial_coverage"
output: html_document
---

## 1. Lib
```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
```

## Data

### 2a. Read data
```{r}
dat_sed <- readxl::read_excel("Input_data/Summary_2019_11_19/sediment_summary.xlsx", sheet = 2, guess_max = 10000)
dat_bio <- readxl::read_excel("Input_data/Summary_2019_11_19/biota_summary.xlsx", sheet = 2, guess_max = 10000)

# xtabs(~shape + colour, dat_sed)
# xtabs(~region, dat_bio) %>% names() %>% dput()
# xtabs(~detGroup, dat_bio) %>% names() %>% dput()
# xtabs(~detGroup, dat_sed) %>% names() %>% dput()

reg_ordered <- c("Barents Sea", "Greenland-Scotland ridge", "Norwegian Sea", 
  "Northern North Sea", "Southern North Sea", "Norwegian Trench", "Skagerrak and Kattegat", "Channel", 
  "Irish and Scottish West Coast", "Irish Sea", "Celtic Sea", 
  "Gulf of Cadiz", "Iberian Sea", "Northern Bay of Biscay")


detgroup_ordered <- c("Metals", "Chlorobiphenyls", "PAH (parent)", "Organo-bromines", 
                      "Organo-metals", "Imposex", "PAH (alkylated)", "Biological effects", 
                      "Dioxins", "Perfluoronated compounds", "Pesticides")

dat_sed$region <- factor(dat_sed$region, levels = reg_ordered)
dat_sed$detGroup <- factor(dat_sed$detGroup, levels = detgroup_ordered)

dat_bio$region <- factor(dat_bio$region, levels = reg_ordered)
dat_bio$detGroup <- factor(dat_bio$detGroup, levels = detgroup_ordered)


```

### 2b. Check new Celtic Sea region
```{r}

# For first check (also in Cornwall)
if (FALSE){
  df <- dat_sed %>%
    filter(determinand == "PB") %>%
    filter(region == "Channel") %>%
    mutate(Popup = paste(longitude, "<BR>", latitude, "<BR>", station, "<BR>", stationName))
}


df <- dat_sed %>%
  filter(determinand == "PB") %>%
  filter(region == "Channel" & latitude < 48.88 & longitude < -1.512) %>%
  mutate(Popup = paste(longitude, "<BR>", latitude, "<BR>", station, "<BR>", stationName))

df2 <- dat_sed %>%
  filter(determinand == "PB") %>%
  filter(region == "Northern Bay of Biscay") %>%
  mutate(Popup = paste(longitude, "<BR>", latitude, "<BR>", station, "<BR>", stationName))

df2a <- df2 %>%
  filter(latitude > 48)
df2b <- df2 %>%
  filter(latitude <= 48)

# Norway (coastline) in UTM coordinates
nc_innerborders <- st_read("Input_data/ospar_inner_boundary_2016_01_002/ospar_inner_boundary_2016_01_002.shp")

# Leaflet map
leaflet(nc_innerborders) %>%
  addTiles() %>%
  addPolylines() %>%
  addMarkers(lng = df$longitude, lat = df$latitude, popup = df$Popup) %>%
  addCircleMarkers(lng = df2a$longitude, lat = df2a$latitude, popup = df2a$Popup) %>%
  addCircleMarkers(lng = df2b$longitude, lat = df2b$latitude, popup = df2b$Popup, color = "red")


```


### 2c. Correct for new Celtic Sea region
```{r}

sel1 <- with(dat_sed, region == "Channel" & latitude < 48.88 & longitude < -1.512); sum(sel1)
sel2 <- with(dat_sed, region == "Northern Bay of Biscay" & latitude > 48.0); sum(sel2)
dat_sed$OSPARregion[sel1 | sel2] <- 3
dat_sed$region[sel1 | sel2] <- "Celtic Sea"

sel1 <- with(dat_bio, region == "Channel" & latitude < 48.88 & longitude < -1.512); sum(sel1)
sel2 <- with(dat_bio, region == "Northern Bay of Biscay" & latitude > 48.0); sum(sel2)
dat_bio$OSPARregion[sel1 | sel2] <- 3
dat_bio$region[sel1 | sel2] <- "Celtic Sea"

```

### 2d. Final region check
```{r}
my_map <- map_data("world")

gg <- dat_sed %>%
  filter(determinand == "PB") %>%
  ggplot(aes(longitude, latitude, color = region)) +
  annotation_map(my_map, color = "lightblue", fill = NA) +
  geom_point() +
  coord_quickmap() +
  theme_void() +
  labs(title = "Region check (lead in sediment)")

ggsave("Figures/Region check 1.png", gg)

gg <- dat_bio %>%
  filter(determinand == "PB") %>%
  ggplot(aes(longitude, latitude, color = region)) +
  annotation_map(my_map, color = "lightblue", fill = NA) +
  geom_point() +
  coord_quickmap() +
  theme_void() +
  labs(title = "Region check (lead in biota)")

ggsave("Figures/Region check 2.png", gg)

gg

```


## 3. Sediment

### a. Tabulate number of stations including informal assessment    
```{r}
df1 <- dat_sed %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  count(OSPARregion, region, detGroup, determinand) 
# df1 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df2 <- df1 %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(n))
# df2 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df_sed_all <- df2 %>%
  spread(detGroup, N_stations_best)

df_sed_all
```

### b. Stations spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_sed_10km <- dat_sed %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  mutate(lat_r = 0.1*floor(latitude*10), 
         lon_r = 0.2*floor(longitude*5),
         Type = case_when(shape %in% "small_open_circle" ~ "Informal status",
                          shape %in% "small_filled_circle" ~ "Status",
                          TRUE ~ "Status and trend"),
         Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, lat_r, lon_r, detGroup, determinand, Type)

#
# Statistics for number of stations per square (and determinand)
#
df <- df_sed_10km %>%
  group_by(lat_r, lon_r, detGroup, determinand) %>%
  summarise(n = sum(n)) 
perc <- 100*table(df$n)/sum(table(df$n))
perc
cumsum(perc)

```

### c. Map 10 km squares 
```{r}
my_map <- map_data("world")

gg <- df_sed_10km %>%
  filter(determinand == "PB") %>%
  ggplot(aes(lon_r, lat_r, color = Type)) +
  annotation_map(my_map, color = "lightblue", fill = NA) +
  geom_point() +
  coord_quickmap() +
  theme_void() +
  labs(title = "Sediment") +
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = rel(1.3)))


ggsave("Figures/02_Sediment_map_squares.png", gg, width = 7, height = 7, dpi = 500)

gg

```

### d. Summarise and graph  
```{r, fig.width=9, fig.height=7}
df_sed_all <- df_sed_10km %>%
  group_by(OSPARregion, region, Type, detGroup, determinand) %>%
  summarise(n_per_det = n()) %>%
  group_by(OSPARregion, region, Type, detGroup) %>%
  summarise(n = n()) %>%
  ungroup()

# For plot 
df <- df_sed_all
sel <- levels(df$region) == "Irish and Scottish West Coast"
levels(df$region)[sel] <- "Irish and Scottish\nWest Coast"
levels(df$region)

gg <- ggplot(df, aes(x = detGroup, y = n, fill = Type)) +
  geom_col(color = "black", size = 0.2) +
  scale_fill_manual(values = c("pink", "green", "blue3")) +
  facet_wrap(vars(region)) +
  theme(axis.text.x = element_text(hjust = 0, angle = -45)) + 
  labs(title = "Sediment (number of grid squares)", y = "Number of 0.1 x 0.2 degree grid squares") +
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = rel(1.2)))

ggsave("Figures/02_Sediment_no_squares.png", gg, width = 6.5, height = 5.5, dpi = 500)

gg

```

### e. Summary table
```{r}
df_sed_summary <- df_sed_all %>% 
  spread(Type, n) %>%
  mutate(Summary = paste(`Informal status`, "/", Status, "/", `Status and trend`)) %>%
  select(OSPARregion, region, detGroup, Summary) %>%
  spread(detGroup, Summary)

```



## 4. Biota

### a. Tabulate number of stations including informal assessment    
```{r}
df1 <- dat_bio %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  count(OSPARregion, region, detGroup, determinand) 
# df1 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df2 <- df1 %>%
  group_by(OSPARregion, region, detGroup) %>%
  summarize(N_stations_best = max(n))
# df2 %>% filter(detGroup %in% "Organo-bromines" & region %in% "Skagerrak and Kattegat")

df_bio_all <- df2 %>%
  spread(detGroup, N_stations_best)

df_bio_all
```

### b. Stations spatially lumped   
Stations < 0.1 lat and 0.2 long apart (order of <10 km apart) lumped      
```{r}
df_bio_10km <- dat_bio %>%
  filter(!detGroup %in% "PAH (alkylated)") %>%
  mutate(lat_r = 0.1*floor(latitude*10), 
         lon_r = 0.2*floor(longitude*5),
         Type = case_when(shape %in% "small_open_circle" ~ "Informal status",
                          shape %in% "small_filled_circle" ~ "Status",
                          TRUE ~ "Status and trend"),
         Location = paste(lat_r, lon_r, sep = "_")) %>%
  count(OSPARregion, region, lat_r, lon_r, detGroup, determinand, Type)

#
# Statistics for number of stations per square (and determinand)
#
df <- df_bio_10km %>%
  group_by(lat_r, lon_r, detGroup, determinand) %>%
  summarise(n = sum(n)) 
perc <- 100*table(df$n)/sum(table(df$n))
perc
cumsum(perc)

```

### c. Map 10 km squares 
```{r}
my_map <- map_data("world")

gg <- df_bio_10km %>%
  filter(determinand == "PB") %>%
  ggplot(aes(lon_r, lat_r, color = Type)) +
  annotation_map(my_map, color = "lightblue", fill = NA) +
  geom_point() +
  coord_quickmap() +
  theme_void() +
  labs(title = "Biota") +
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = rel(1.3)))

ggsave("Figures/02_Biota_map_squares.png", gg, width = 7, height = 7, dpi = 500)

gg

```

### d. Summarise and graph  
```{r, fig.width=9, fig.height=7}
df_bio_all <- df_bio_10km %>%
  group_by(OSPARregion, region, Type, detGroup, determinand) %>%
  summarise(n_per_det = n()) %>%
  group_by(OSPARregion, region, Type, detGroup) %>%
  summarise(n = n())

# For plot 
df <- df_bio_all
sel <- levels(df$region) == "Irish and Scottish West Coast"
levels(df$region)[sel] <- "Irish and Scottish\nWest Coast"
levels(df$region)

gg <- ggplot(df, aes(x = detGroup, y = n, fill = Type)) +
  geom_col(color = "black", size = 0.2) +
  scale_fill_manual(values = c("pink", "green", "blue3")) +
  facet_wrap(vars(region)) +
  theme(axis.text.x = element_text(hjust = 0, angle = -45)) + 
  labs(title = "Biota (number of grid squares)", y = "Number of 0.1 x 0.2 degree grid squares") +
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = rel(1.2)))

ggsave("Figures/02_Biota_no_squares.png", gg, width = 8, height = 5.5, dpi = 500)

gg

```

### e. Summary table
```{r}
df_bio_summary <- df_bio_all %>% 
  spread(Type, n) %>%
  mutate(Summary = paste(`Informal status`, "/", Status, "/", `Status and trend`)) %>%
  select(OSPARregion, region, detGroup, Summary) %>%
  spread(detGroup, Summary)

```

## 5. Saving  
### a. Add missing regions in proper order
```{r}
df_regions <- dat_bio %>% count(OSPARregion, region) %>% select(-n)

df_sed_all <- df_regions %>%
  full_join(df_sed_all) 

df_sed_summary <- df_regions %>%
  full_join(df_sed_summary) 

df_bio_all <- df_regions %>%
  full_join(df_bio_all) 

df_bio_summary <- df_regions %>%
  full_join(df_bio_summary) 


# ?full_join
```

### b. Save as excel

```{r}
info <- tibble(
  `Numbers in summary tables: number of 10 km squares with` = 
    "Informal status / Status / Status and trend"
)

df_list <- list(
  df_sed_all, df_sed_summary,
  df_bio_all, df_bio_summary, 
  info
)

names(df_list) <- c(
  "sed_all", "sed_summary", 
  "bio_all", "bio_summary",
  "Info"
)

openxlsx::write.xlsx(df_list, "Produced_data/02_Spatial_coverage_2019.xlsx")


  # ?extract

```

