---
title: "03_Check_Norwegian_data"
author: "DHJ"
date: "17 12 2019"
output: html_document
---

## 1. Libs
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)

source("03_Check_Norwegian_data_spec.R", encoding = "UTF-8")

```

## 2. Read MILKYS data  
```{r}
dat <- readRDS("../Milkys/Data/10_data_all.RData")

dat01 <- readRDS("../Milkys/Data/01_dat_all.rds")

str(dat)

# Not a good result
# datxl <- readxl::read_excel("Input_data/Trendtabell_20160919_Komplett_v4_GSE med samletabell.xlsx")
datxl <- readr::read_csv2(
  "Input_data/Trendtabell_20160919_Komplett_v4_GSE med samletabell.csv",
  col_types = spec_bigexcel
  )

# spec(datxl)

```

## 3. Check up on the 98B stations  
After questions from Rob
```{r}
dat %>% 
  filter(substr(STATION_CODE,1,3) %in% "98B") %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

```

## 4. Check up on some other stations not marked 
After questions from Rob   

### 45B  
- 10 years of data
- Seems clear that 45B2 (Hammerfest harbour area) actually differs from the others.  
- Perhaps also 45B (Hammerfest area) differs from 45B1  
```{r}
short_code <- "45B"
no_letters <- nchar(short_code)

# Number of years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR) %>%
  nrow()  

# Species and tissues
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(LATIN_NAME, TISSUE_NAME, STATION_CODE)

# Years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

# Check percentages of different PCBs
tb <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(substr(PARAM,1,2) %in% "CB") %>%
  group_by(MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)
tb2 <- tb <- tb[,1:10]
for (i in 3:10) tb2[,i] <- round(tb2[,i]/tb2[,2]*100, 1)
tb2

# Select some parameters
df <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(PARAM %in% c("CD", "HG", "CB101", "CB180"))

# Show medians
df %>%
  group_by(TISSUE_NAME, MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)

# Show medians, graph
df %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  ggplot(aes(MYEAR, Conc, color = STATION_CODE)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(vars(TISSUE_NAME, PARAM), scales = "free_y")
  # facet_grid(rows = vars(TISSUE_NAME), cols = vars(PARAM), scales = "free_y")



```



### 43F  
- Only on year of data (1996)  
- But the old excel table ("Trendtabell_20160919_Komplett_v4_GSE med samletabell.xlsx") has no 43F, but has 43F1 in 2006-07?!  
```{r}
short_code <- "43F"
no_letters <- nchar(short_code)

# Number of years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR) %>%
  nrow()  

# Species and tissues
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(LATIN_NAME, TISSUE_NAME, STATION_CODE)

# Years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

# Check percentages of different PCBs
tb <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(substr(PARAM,1,2) %in% "CB") %>%
  group_by(MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)
tb2 <- tb <- tb[,1:10]
for (i in 3:10) tb2[,i] <- round(tb2[,i]/tb2[,2]*100, 1)
tb2

# Select some parameters
df <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(PARAM %in% c("CD", "HG", "CB101", "CB180"))

# Show medians
df %>%
  group_by(TISSUE_NAME, MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)

# Show medians, graph
df %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  ggplot(aes(MYEAR, Conc, color = STATION_CODE)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(vars(TISSUE_NAME, PARAM), scales = "free_y")
  # facet_grid(rows = vars(TISSUE_NAME), cols = vars(PARAM), scales = "free_y")

```



### 98F    
- Find only 98F, 4 years  
```{r}

short_code <- "98F"
no_letters <- nchar(short_code)

# Number of years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR) %>%
  nrow()  

# Species and tissues
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(LATIN_NAME, TISSUE_NAME, STATION_CODE)

# Years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

# Check percentages of different PCBs
tb <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(substr(PARAM,1,2) %in% "CB") %>%
  group_by(MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)
tb2 <- tb <- tb[,1:10]
for (i in 3:10) tb2[,i] <- round(tb2[,i]/tb2[,2]*100, 1)
tb2

# Select some parameters
df <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(PARAM %in% c("CD", "HG", "CB101", "CB180"))

# Show medians
df %>%
  group_by(TISSUE_NAME, MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)

# Show medians, graph
df %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  ggplot(aes(MYEAR, Conc, color = STATION_CODE)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(vars(TISSUE_NAME, PARAM), scales = "free_y")
  # facet_grid(rows = vars(TISSUE_NAME), cols = vars(PARAM), scales = "free_y")

```


### 92A  
92A1 1992-1997, 92A2 2005 
```{r}
short_code <- "92A"
no_letters <- nchar(short_code)

# Species and tissues
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(LATIN_NAME, TISSUE_NAME, STATION_CODE)

# Years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

# Check percentages of different PCBs
tb <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(substr(PARAM,1,2) %in% "CB") %>%
  group_by(MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)
tb2 <- tb <- tb[,1:10]
for (i in 3:10) tb2[,i] <- round(tb2[,i]/tb2[,2]*100, 1)
tb2

# Select some parameters
df <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(PARAM %in% c("CD", "HG", "CB101", "CB180"))

# Show medians
df %>%
  group_by(TISSUE_NAME, MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)

# Show medians, graph
df %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  ggplot(aes(MYEAR, Conc, color = STATION_CODE)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(vars(TISSUE_NAME, PARAM), scales = "free_y")
  # facet_grid(rows = vars(TISSUE_NAME), cols = vars(PARAM), scales = "free_y")

```

### 227G  
92A1 1992-1997, 92A2 2005 
```{r}
short_code <- "227G"
no_letters <- nchar(short_code)

# Species and tissues
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(LATIN_NAME, TISSUE_NAME, STATION_CODE)

# Years
dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  count(MYEAR, STATION_CODE) %>%
  pivot_wider(names_from = STATION_CODE, values_from = n)

# Check percentages of different PCBs
tb <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(substr(PARAM,1,2) %in% "CB") %>%
  group_by(MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)
tb2 <- tb <- tb[,1:10]
for (i in 3:10) tb2[,i] <- round(tb2[,i]/tb2[,2]*100, 1)
tb2

# Select some parameters
df <- dat %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  filter(PARAM %in% c("CD", "HG", "CB101", "CB180"))

# Show medians
df %>%
  group_by(TISSUE_NAME, MYEAR, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  pivot_wider(names_from = PARAM, values_from = Conc)

# Show medians, graph
df %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, PARAM) %>%
  summarize(Conc = median(VALUE_WW)) %>%
  ggplot(aes(MYEAR, Conc, color = STATION_CODE)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(vars(TISSUE_NAME, PARAM), scales = "free_y")
  # facet_grid(rows = vars(TISSUE_NAME), cols = vars(PARAM), scales = "free_y")

```


```{r}

summarize_sequence <- function(x){
  x <- sort(unique(x))
  dx <- diff(x)
  df <- tibble(
    x = x,
    index = cumsum(c(1, dx) > 1) + 1)
  df %>% 
    group_by(index) %>%
    summarize(Min = min(x),Max = max(x)) %>%
    mutate(Summ = ifelse(Min < Max, paste0(Min,"-",Max), Min)) %>%
    ungroup() %>%
    summarize(Summ = paste0(Summ, collapse = ",")) %>%
    pull(Summ)
}

# summarize_sequence(c(1,2,5,6,7,9,10,11,12))


```


## Compare station codes in database and in the big Excel file of Tore (THO)

### Example: 227G  
92A1 1992-1997, 92A2 2005 
```{r}

short_code <- "227G"
no_letters <- nchar(short_code)

# Stations and years in the database
data_yrs <- dat01 %>% 
  filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
  group_by(MYEAR, STATION_CODE) %>%
  summarise(n_data = n())

# Big excel file - confirm unique records
datxl %>%
  filter(substr(`Station Code`, 1, no_letters) %in% short_code) %>%
  count(`Parameter Code`, `Station Code`, `Species`, `Tissue`, `Basis`) %>%
  count(n)
  
# Big excel file - put on long format
datxl_long <- datxl %>%
  filter(substr(`Station Code`, 1, no_letters) %in% short_code) %>%
  select(`Parameter Code`, `Station Code`, `Species`, `Tissue`, `Basis`, `1980`:`2015`) %>%
  pivot_longer(cols = `1980`:`2015`, names_to = "MYEAR", values_to = "VALUE") %>%
  mutate(MYEAR = as.numeric(MYEAR)) %>%
  rename(STATION_CODE = `Station Code`)

# Stations and years in the big excel file
xlfile_yrs <- datxl_long %>%
  filter(VALUE != "") %>%
  group_by(MYEAR, STATION_CODE) %>%
  summarise(n_xl = n())

# Stations and years in the database and in the big excel file, combined
compare_yrs <- full_join(data_yrs, xlfile_yrs) %>%
  ungroup()

# Compare
compare_yrs %>%
  group_by(STATION_CODE) %>%
  summarise(
    database_only = summarize_sequence(MYEAR[n_data > 0 & is.na(n_xl)]),
    both = summarize_sequence(MYEAR[!is.na(n_data) & !is.na(n_xl)]),
    excel_only = summarize_sequence(MYEAR[is.na(n_data) & n_xl > 0])
    )


```

### Function  
```{r}
check_stations <- function(short_code){
  no_letters <- nchar(short_code)
  
  # Stations and years in the database
  data_yrs <- dat01 %>% 
    filter(substr(STATION_CODE,1, no_letters) %in% short_code) %>%
    group_by(MYEAR, STATION_CODE) %>%
    summarise(n_data = n())
  
  # Big excel file - put on long format
  datxl_long <- datxl %>%
    filter(substr(`Station Code`, 1, no_letters) %in% short_code) %>%
    select(`Parameter Code`, `Station Code`, `Species`, `Tissue`, `Basis`, `1980`:`2015`) %>%
    pivot_longer(cols = `1980`:`2015`, names_to = "MYEAR", values_to = "VALUE") %>%
    mutate(MYEAR = as.numeric(MYEAR)) %>%
    rename(STATION_CODE = `Station Code`)
  
  # Stations and years in the big excel file
  xlfile_yrs <- datxl_long %>%
    filter(VALUE != "") %>%
    group_by(MYEAR, STATION_CODE) %>%
    summarise(n_xl = n())
  
  # Stations and years in the database and in the big excel file, combined
  compare_yrs <- full_join(data_yrs, xlfile_yrs, by = c("MYEAR", "STATION_CODE")) %>%
    ungroup()
  
  # Compare
  compare_yrs %>%
    group_by(STATION_CODE) %>%
    summarise(
      database_only = summarize_sequence(MYEAR[n_data > 0 & is.na(n_xl)]),
      both = summarize_sequence(MYEAR[!is.na(n_data) & !is.na(n_xl)]),
      excel_only = summarize_sequence(MYEAR[is.na(n_data) & n_xl > 0])
    )
  
}

# check_stations("227G")
```

### Checking
```{r}
check_stations("45B")
check_stations("43F")
check_stations("98F")
check_stations("92A")
check_stations("227G")

```


