---
title: "05: Check CEMP Appendices"
author: "DHJ"
date: "27 2 2020"
output: html_document
---

For different groups of chemicals, check first year of data, in order to set as *Temporal scope* first year  

## 1. Packages and functions  
```{r}

library(plyr)
library(dplyr)
library(purrr)

source('../Milkys/00_read_ICES_webservice_functions.R', encoding = 'ANSI_X3.4-1986')

# Downloaded from https://vocab.ices.dk/?ref=22
df_countries <- read.csv('../Milkys/00_read_ICES_webservice_functions_countries.csv', encoding = 'UTF-8')

# Downloaded from 
df_parametergroups <- read.csv('../Milkys/00_read_ICES_webservice_functions_groups.csv', encoding = 'UTF-8')


```

```{r}
"http://dome.ices.dk/Webservices/DOMEWebServices.asmx/selectContaminantsInSediment?PARAM=CD&RLABO=ALUK&ALABO=&yearBegining=&yearEnd=&MATRX=&TAXA=&PURPM=&MPROG=&Area=&CNTRY=&ParamGroup="
```


## 2. Countrycodes and function
```{r}

df_countries <- df_countries %>%
  filter(Description %in% c("Iceland", "Norway", "Denmark", "Sweden", "Germany", "The Netherlands", "Belgium",
                            "United Kingdom", "Ireland", "France", "Spain", "Portugal")
  
)

get_ices_biotadata_polite <- function(...){
  result <- get_ices_biotadata(...)
  Sys.sleep(1)
  result
}
# Test 1
# df <- get_ices_biotadata_polite("TBTIN", yearstart = 2010, yearend = 2018, country = 74)
#	Test 2: Get all data for a parameter group, here: dioxins (OC-DX) in Norway
# df <- get_ices_biotadata_polite(
#   "", yearstart = 1990, yearend = 2000, country = 58, paramgroup = "OC-DX")

#
# Get biota data from several countries:
#
get_ices_biotadata_all <- function(param, countrycodes,
                                   yearstart = NULL, yearend = yearstart, 
                                   country = "", matrix = "", lab = "", species = "",
                                   paramgroup = ""){
  codes <- countrycodes
  names(codes) <- countrycodes
  codes %>% 
    map_df(~get_ices_biotadata_polite(param, country = as.character(.), 
                               yearstart = yearstart, yearend = yearend, matrix = matrix,
                               lab = lab, species = species, paramgroup = paramgroup),
           .id = "Code")
}

# Tests
 test1 <- get_ices_biotadata_all("TBTIN", countrycodes = c(74,77), yearstart = 2010, yearend = 2018)  # UK,  test1 <- get_ices_biotadata_all("TBTIN", countrycodes = c(74,77), yearstart = 2010, yearend = 2018)  # UK, Sweden
# test2 <- get_ices_biotadata_all("CB118", countrycodes = df_countries$Code, yearstart = 2017, yearend = 2019)  # All countries

get_ices_sedimentdata_polite <- function(...){
  result <- get_ices_sedimentdata(...)
  Sys.sleep(1)
  result
}
get_ices_sedimentdata_polite("CD", yearstart = 1996, country = 74)

#
# Get sediment data from several countries:
#
get_ices_sedimentdata_all <- function(param, countrycodes,
                                   yearstart = NULL, yearend = yearstart, 
                                   country = "", lab = "", 
                                   paramgroup = ""){
  codes <- countrycodes
  names(codes) <- countrycodes
  codes %>% 
    map_df(~get_ices_sedimentdata(param, country = as.character(.), 
                               yearstart = yearstart, yearend = yearend, 
                               lab = lab, paramgroup = paramgroup),
           .id = "Code")
}
# test3 <- get_ices_sedimentdata_all("CD", yearstart = 1996, countrycodes = df_countries$Code)    
# xtabs(~Country, test3)
 
```

## 3a. Appendix H6: Planar PCBs in biota       
- CB15,  CB37,  CB77,  CB81, CB126 , CB169

```{r, results='hide', message=FALSE}
df_ices <- c("CB15",  "CB37",  "CB77",  "CB81", "CB126", "CB169") %>%
  map_df(~get_ices_biotadata_all(., countrycodes = df_countries$Code, yearstart = 1970, yearend = 2019))  # UK, Sweden
table(df_ices$Code)

```

#### Result
```{r}

xtabs(~MYEAR + PARAM, df_ices)
xtabs(~MYEAR + Country + grepl("CEMP", MPROG), df_ices)

```


## 3b. Appendix H6: Check planar PCBs in sediment   
- CB15,  CB37,  CB77,  CB81, CB126 , CB169  
```{r}

df_ices <- c("CB15",  "CB37",  "CB77",  "CB81", "CB126", "CB169") %>%
  map_df(~get_ices_sedimentdata_all(., countrycodes = df_countries$Code, yearstart = 1970, yearend = 2019))  # UK, Sweden
table(df_ices$Code)

```

#### Result
```{r}
xtabs(~MYEAR + PARAM, df_ices)
xtabs(~MYEAR + Country + grepl("CEMP", MPROG), df_ices)

```

## 4. Appendix H9: Check PCDF in sediment and biota    
I.e., Polychlorinated dibenzodioxins and furans   
E.g., 
- CDD1N =	1,2,3,7,8-pentachlorodibenzo-p-dioxin     
- CDD9X = 1,2,3,7,8,9-hexachlorodibenzo-p-dioxin   
- CDF2T = 2,3,7,8-tetrachloro-dibenzofuran
   
Earliest data are from **1995**  
```{r}
df_dioxins <- get_ices_biotadata_all("", countrycodes = df_countries$Code, 
                                     yearstart = 1970, yearend = 2019,
                                     paramgroup = "OC-DX")  # see df_parametergroups

table(df_dioxins$Code)
xtabs(~MYEAR + PARAM, df_dioxins)
xtabs(~MYEAR + Country + grepl("CEMP", MPROG), df_dioxins)

```

## 5. Appendix H10: PAH and metal-specific relevant Biological Effects   

- Includes PAH metabolites + ALA-D, and EROD ("the EROD catalytic enzyme assay is the technique recommended for monitoring CYP 1A activity")  
- Probably NOT on this group:  
    * %HCT = hematocrit - volume percentage of red blood cells in blood   
    * CAT = catalase activity (Unit example: micromol/min/mg protein)   
   
Earliest year = **1989** (EROD in Sweden)    
```{r}
df_biol <- get_ices_biotadata_all("", countrycodes = df_countries$Code, 
                                     yearstart = 1970, yearend = 2019,
                                     paramgroup = "B-MBA")  # see df_parametergroups

table(df_biol$Code)
xtabs(~MYEAR + PARAM, df_biol %>% filter(grepl("CEMP", MPROG)))
xtabs(~MYEAR + Country + grepl("CEMP", MPROG), df_biol)

xtabs(~MYEAR + grepl("CEMP", MPROG), df_biol)

```