---
title: "Time trends in blue mussel, comparing wet-weight and dry weight concentrations"
author: "DHJ"
date: "24 11 2020"
output: 
  html_document:
    toc: true
    toc_float: true
  
---


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)
library(plotly)
library(htmlTable)

setHtmlTableTheme(css.rgroup = "font-weight: normal",
                  css.cgroup = "font-weight: normal",
                  css.header = "font-weight: normal")


```

```{r, echo = FALSE}

#
# Functions for Deming (orthogonal) regression
# a.k.a total sum-of-squares regression
#   (i.e. it is assumed that there are errors in both x and y)  
# By default, the noise_ratio (the ratio between y and x errors) is 
#   assumed to equal sd(y)/sd(x)  
#
# From
# https://stackoverflow.com/a/59172951/1734247
#

deming.fit <- function(x, y, noise_ratio = sd(y)/sd(x)) {
  if(missing(noise_ratio) || is.null(noise_ratio)) noise_ratio <- eval(formals(sys.function(0))$noise_ratio) # this is just a complicated way to write `sd(y)/sd(x)`
  delta <-  noise_ratio^2
  x_name <- deparse(substitute(x))

  s_yy <- var(y)
  s_xx <- var(x)
  s_xy <- cov(x, y)
  beta1 <- (s_yy - delta*s_xx + sqrt((s_yy - delta*s_xx)^2 + 4*delta*s_xy^2)) / (2*s_xy)
  beta0 <- mean(y) - beta1 * mean(x) 

  res <- c(beta0 = beta0, beta1 = beta1)
  names(res) <- c("(Intercept)", x_name)
  class(res) <- "Deming"
  res
}

deming <- function(formula, data, R = 100, noise_ratio = NULL, ...){
  ret <- boot::boot(
    data = model.frame(formula, data), 
    statistic = function(data, ind) {
      data <- data[ind, ]
      args <- rlang::parse_exprs(colnames(data))
      names(args) <- c("y", "x")
      rlang::eval_tidy(rlang::expr(deming.fit(!!!args, noise_ratio = noise_ratio)), data, env = rlang::current_env())
    },
    R=R
  )
  class(ret) <- c("Deming", class(ret))
  ret  
}

predictdf.Deming <- function(model, xseq, se, level) {
  pred <- as.vector(tcrossprod(model$t0, cbind(1, xseq)))
  if(se) {
    preds <- tcrossprod(model$t, cbind(1, xseq))
    data.frame(
      x = xseq,
      y = pred,
      ymin = apply(preds, 2, function(x) quantile(x, probs = (1-level)/2)),
      ymax = apply(preds, 2, function(x) quantile(x, probs = 1-((1-level)/2)))
    )
  } else {
    return(data.frame(x = xseq, y = pred))
  }
}


```


```{r, echo=FALSE}

df1 <- readRDS("OSPAR_MIME_2020/shellfish_ww_2020.rds")
df2 <- readRDS("OSPAR_MIME_2020/shellfish_dw_2021.rds")


# names(df1) 
# names(df2)
# names(df1) %>% paste(collapse = ", ")


#
# Check which variables defines unique blue mussel records 
#
if (FALSE){

df1 %>%
  filter(species == "Mytilus edulis") %>%
  distinct(region, subregion, ICES_ecoregion, country, station_code, station_name, station_long_name, 
           latitude, longitude, MSTAT, WLTYP, determinand, determinand_group, species, 
           matrix, basis, unit, sex, metoa, mammal_group) %>%
  nrow()

df1 %>%
  filter(species == "Mytilus edulis") %>%
  distinct(country, station_code, determinand, matrix, basis, unit) %>%
  nrow()

}

```

```{r, echo=FALSE}

## Join  

### Change names of regression variables

# names(df1) %>% dput()

vars <- c(
  "n_year_all", "n_year_fit", "n_year_positive", "first_year_all", 
  "first_year_fit", "last_year", "p_nonlinear", "p_linear", "p_overall", 
  "p_linear_trend", "linear_trend", "p_recent_trend", "recent_trend", 
  "detectable_trend", "mean_last_year", "climit_last_year", "BAC", 
  "BAC_diff", "BAC_achieved", "BAC_below", "EAC", "EAC_diff", "EAC_achieved", 
  "EAC_below", "MPC", "MPC_diff", "MPC_achieved", "MPC_below", 
  "FEQG", "FEQG_diff", "FEQG_achieved", "FEQG_below", "imposex_class")

for (var in vars){
  names(df1)[names(df1) == var] <- paste0(var, "_ww")
  names(df2)[names(df2) == var] <- paste0(var, "_dw")
}


```

```{r, echo=FALSE, results='hide', message=FALSE}

### The actual join  

# names(df1) %>% dput()

df <- df1 %>% 
  filter(species == "Mytilus edulis") %>%
  select(region, subregion, ICES_ecoregion, country, station_code, station_name, station_long_name, 
           latitude, longitude, MSTAT, WLTYP, determinand, determinand_group, species, 
           matrix, basis, unit, sex, metoa, mammal_group,
         n_year_all_ww:imposex_class_ww) %>%
  left_join(
    df2 %>%
      filter(species == "Mytilus edulis") %>%
      select(country, station_code, determinand,   # matrix, basis, unit,
           n_year_all_dw:imposex_class_dw)
  )

nrow(df1)
nrow(filter(df1, species == "Mytilus edulis"))
nrow(df)

```


## Statistics   
All statistics are for blue mussel   

### Overview     
Number of blue mussel time series per country    
```{r, echo=FALSE}

df %>%
  filter(species == "Mytilus edulis") %>% # View()
  xtabs(~ determinand_group + country, .)

```

### Number of significant linear trends   
For metals and chlorobiphenyls separately 
```{r, echo=FALSE, results = 'asis'}

for (group in c("Metals", "Chlorobiphenyls", "Polybrominated diphenyl ethers", 
                "PAH parent compounds", "Organochlorines (other)")){
  df %>%
    filter(species == "Mytilus edulis" & 
             determinand_group %in% group) %>% # View()
    xtabs(~ (p_linear_trend_ww <= 0.05) + (p_linear_trend_dw <= 0.05), .) %>%
    htmlTable(caption = group) %>%
    print()
}


```


### Trends, up & down 
```{r, echo=FALSE}

trend_order <- c(
  "declining (ww & dw)", "declining (ww only)", "declining (dw only)", 
  "increasing (ww & dw)", "increasing (ww only)", "increasing (dw only)", 
  "declining ww, increasing dw", "increasing ww, declining dw", 
  "no trend (ww & dw)")

df %>%
  filter(species == "Mytilus edulis" &
           determinand_group %in% c("Metals", "Chlorobiphenyls", "Polybrominated diphenyl ethers", 
                "PAH parent compounds", "Organochlorines (other)")) %>% 
  mutate(
    determinand_group = levels(determinand_group)[as.numeric(determinand_group)],
    Trend = case_when(
      p_linear_trend_ww <= 0.05 & linear_trend_ww < 0 & p_linear_trend_dw <= 0.05 & linear_trend_dw < 0 ~ "declining (ww & dw)", 
      p_linear_trend_ww <= 0.05 & linear_trend_ww < 0 & p_linear_trend_dw > 0.05 ~ "declining (ww only)", 
      p_linear_trend_dw <= 0.05 & linear_trend_dw < 0 & p_linear_trend_ww > 0.05 ~ "declining (dw only)", 
      p_linear_trend_ww <= 0.05 & linear_trend_ww > 0 & p_linear_trend_dw <= 0.05 & linear_trend_dw > 0 ~ "increasing (ww & dw)", 
      p_linear_trend_ww <= 0.05 & linear_trend_ww > 0 & p_linear_trend_dw > 0.05 ~ "increasing (ww only)", 
      p_linear_trend_dw <= 0.05 & linear_trend_dw > 0 & p_linear_trend_ww > 0.05~ "increasing (dw only)", 
      p_linear_trend_ww <= 0.05 & linear_trend_ww < 0 & p_linear_trend_dw <= 0.05 & linear_trend_dw > 0 ~ "declining ww, increasing dw", 
      p_linear_trend_ww <= 0.05 & linear_trend_ww > 0 & p_linear_trend_dw <= 0.05 & linear_trend_dw < 0 ~ "increasing ww, declining dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "no trend (ww & dw)"),
    Trend = factor(Trend, levels = trend_order)
  ) %>%
  xtabs(~ Trend + determinand_group, .) %>%
  htmlTable()

```




## Plots of slope of linear trends   

X axis: Slope of time trend (i.e., change per year), wet-weight concentration   
Y axis: Slope of time trend (i.e., change per year), wet-weight concentration   

* Red line: Regression line (red) calculated using Deming (orthogonal) regression   
* Blue dotted line: 1:1 line (same trend using w.w. and d.w. concentration)        

### Overall per group   
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8}

p_levels <- c("ww and dw", "ww", "dw", "none") 
p_colors <- c("red3", "magenta", "dodgerblue", "grey35")
names(p_colors) <- p_levels


df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% c("Metals", "Chlorobiphenyls")) %>% # View()
  mutate(
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw)) +
  geom_point(aes(color = `P < 0.05`)) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  geom_smooth(method = "deming", color = "red", size = 1) +
  facet_wrap(vars(determinand_group))

```


## Individual compounds {.tabset }  


### PCBs
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=7}


gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>% # View()
  add_count(determinand) %>%
  filter(n > 120) %>%
  mutate(
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`)) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand))

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### PCBs (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```

### Metals  
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=7}

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  mutate(
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`)) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand))

gg + geom_smooth(method = "deming", color = "red", size = 1)
 
```

### Metals (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


## Individual compounds, difference plot {.tabset } 

### PCBs
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>% # View()
  add_count(determinand) %>%
  filter(n > 120) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(trend_mean_dw_ww, trend_dw_minus_ww, text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`)) +
  scale_color_manual(values = p_colors) +
  geom_hline(yintercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand))
 
gg 


```

### PCBs (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "lm", color = "red", se = FALSE)

ggplotly(gg2)

```

### Metals
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>% # View()
  add_count(determinand) %>%
  filter(n > 120) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(trend_mean_dw_ww, trend_dw_minus_ww, text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`)) +
  scale_color_manual(values = p_colors) +
  geom_hline(yintercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand))
 
gg + geom_smooth(method = "lm", color = "red", size = 1)


```


### Metals (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "lm", color = "red", se = FALSE)

ggplotly(gg2)

```


## PCBs by country  {.tabset}

### UK 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "United Kingdom"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### UK (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```

### Ireland 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Ireland"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)
 
gg + geom_smooth(method = "deming", color = "red", size = 1)

```


### IE (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### France 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "France"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
    mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)


```

### FR (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Spain 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Spain"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)
 
```


### ES (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Iceland 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Iceland"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```


### IS (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```



### Norway 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Norway"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Chlorobiphenyls") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```


### NO (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```



## Metals by country  {.tabset}

### UK 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "United Kingdom"
gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```


### UK (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Ireland 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Ireland"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### IE (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### France 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "France"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### FR (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Spain 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Spain"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)
 
gg + geom_smooth(method = "deming", color = "red", size = 1)

```


### ES (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Denmark 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Denmark"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)
 
```

### DK (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```


### Iceland 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Iceland"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### IS (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + 
  geom_smooth(method = "lm", color = "red", se = FALSE)

ggplotly(gg2)

```


### Norway 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

sel_country <- "Norway"

gg <- df %>%
  filter(species == "Mytilus edulis" & 
           determinand_group %in% "Metals") %>%
  add_count(determinand) %>%
  filter(n > 120) %>%
  filter(country %in% sel_country) %>%
  mutate(
    trend_mean_dw_ww = (linear_trend_dw + linear_trend_ww)/2,
    trend_dw_minus_ww = linear_trend_dw - linear_trend_ww,
    `P < 0.05` = case_when(
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw <= 0.05 ~ "ww and dw", 
      p_linear_trend_ww <= 0.05 & p_linear_trend_dw > 0.05 ~ "ww", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw <= 0.05 ~ "dw", 
      p_linear_trend_ww > 0.05 & p_linear_trend_dw > 0.05 ~ "none"),
    `P < 0.05` = factor(`P < 0.05`, levels = p_levels)
  ) %>%
  ggplot(aes(linear_trend_ww, linear_trend_dw, 
             text = paste(country, station_code, station_name))) +
  geom_point(aes(color = `P < 0.05`), size = 1) +
  scale_color_manual(values = p_colors) +
  geom_abline(slope = 1, intercept = 0, color = "blue", size = 1, linetype = 2) +
  facet_wrap(vars(determinand)) +
  labs(title = sel_country)

gg + geom_smooth(method = "deming", color = "red", size = 1)

```

### NO (interactive)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

gg2 <- gg + geom_smooth(method = "deming", color = "red", se = FALSE)

ggplotly(gg2)

```







