---
title: "Correlations between fat and organics - Norwegian cod"
author: "DHJ"
date: "25 11 2020"
output: 
  html_document:
    toc: true
    toc_float: true
---


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)
# library(plotly)
library(mgcv)
library(purrr)
library(visreg)

```


```{r, echo = FALSE}

#
# Data (Norwegian CEMP data on sample level)
#

df <- readRDS("../Milkys2_pc/Files_from_Jupyterhub_2019/Raw_data/109_adjusted_data_2020-08-05.rds")

# df %>%
#   filter(LATIN_NAME %in% "Gadus morhua" & TISSUE_NAME %in% "Lever") %>%
#   group_by(PARAM, STATION_CODE) %>%
#   summarise(n_year = length(unique(MYEAR)), over_loq = mean(is.na(FLAG1))) %>%
#   filter(n_year > 10 & over_loq > 0.7) %>% pull(PARAM) %>% unique() %>% dput()


```

```{r, echo = FALSE}

param <- "BDE49"
station <- "30B"

```


## Example: one substance (`r param`), one station (`r station`)    
  
Model: `ln(concentration) = a*ln(fat) + b*Length + spline(Year)`  

Thus:  

* The effect of ln(fat) assumed to be linear   
* Effect of year (non-linear effect) and length (linear) also taken into account   
```{r, echo = FALSE}

# param <- "CB_S7"


#
# Test 1 (result not shown): one substance, all stations    
#
df2 <- df %>%
  filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% param & TISSUE_NAME %in% "Lever" & !is.na(FAT_PERC))

# Don't show these results  
if (FALSE){
  summary(m1)
  par(mfrow = c(2,2), mar = c(4,5,2,1), oma = c(0,0,2,0))
  visreg(m1)
  mtext(param, outer = TRUE, line = 0.5)
}


#
# Test 2 (shown): one substance + one station
#

df3_check <- df2 %>%
  filter(STATION_CODE == station)

# All stations    
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df3_check)
summary(m1)

par(mfrow = c(2,2), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

# summary(m1) %>% str()
# summary(m1)$p.table[2,]

```


### Several substances + stations          
```{r, echo = FALSE}

#
# Test with one substance + one station
#

# Test with one substance  

get_estimates <- function(param, station, data){
  
  df2 <- data %>%
    filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% param & 
             TISSUE_NAME %in% "Lever" & 
             STATION_CODE == station & 
             !is.na(FAT_PERC) & !is.na(LNMEA))
  
  m0 <- gam(log(VALUE_WW) ~ 1, data = df2)
  m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + LNMEA, data = df2)
  m2 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)
  dev <- anova(m0, m1, m2)[,2]
  
  tibble(PARAM = param, 
         STATION_CODE = station,
         Fat_slope = summary(m2)$p.table[2,1],
         Fat_slope_se = summary(m2)$p.table[2,2],
         Fat_pvalue = summary(m2)$p.table[2,4],
         Dev_expl_tot = (dev[1]-dev[3])/dev[1],
         Dev_expl_fat = (dev[2]-dev[3])/dev[1]
         )
  
}

# Test
# get_estimates("CB118", "30B", df)

params_all <- c(
  "BDE28", "BDE47", "BDE49", 
  "BDE66", "BDE99", "BDE100", "BDE154", "BDESS", "BDE6S",
  "CB28", "CB52", "CB101", "CB105", 
  "CB118", "CB138", "CB153", "CB156", "CB180", "CB209", "CB_S7", 
  "SCCP", "MCCP", "OCS", "HCB",  
  "DDEPP", "DDTEP", "DDTPP", "TDEPP",    # expected to NOT correlate with fat
  "PFAS", "PFOS", "PFOSA")               # same

#
# Series of PCBs and BDEs  
#
# Keep time series with at least 10 years and at least 70% data over LOQ
#
df_series <- df %>%
  filter(LATIN_NAME %in% "Gadus morhua" & TISSUE_NAME %in% "Lever" & 
           PARAM %in% params_all) %>%
  group_by(PARAM, STATION_CODE) %>%
  summarise(
    n_year = length(unique(MYEAR)), 
    over_loq = mean(is.na(FLAG1)),
    .groups = "drop") %>%
  filter(n_year > 10 & over_loq > 0.7)   # at least 10 years and at least 70% data over LOQ

#
# Run analysis for each time series
#
df_effect <- 1:nrow(df_series) %>% 
  purrr::map_dfr(
    ~get_estimates(param = df_series$PARAM[.], station = df_series$STATION_CODE[.], df)
  )

df_effect <- df_effect %>%
  mutate(Fat_effect = ifelse(Fat_pvalue < 0.05, "P < 0.05",  "P >= 0.05"))

```


## Effect of fat, slope {.tabset}  

Effect of ln(fat) on ln(concentration)   

* Analysis on station-by-station basis   
* The effect of ln(fat) assumed to be linear   
* Effect of year (non-linear effct) and length (linear) also taken into account   
* Analysis done for time series with at least 10 years and at least 70% data over LOQ    
  
### PCBs
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c(
  "CB28", "CB52", "CB101", "CB105", 
  "CB118", "CB138", "CB153", "CB156", "CB180", "CB209", "CB_S7")

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```


### BDEs
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c(
  "BDE28", "BDE47", "BDE49", 
  "BDE66", "BDE99", "BDE100", "BDE154", "BDESS", "BDE6S")


df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```



### Organochorines, paraffins
```{r, echo=FALSE, fig.width=6, fig.height=4}

params <- c("OCS", "HCB", "SCCP", "MCCP" )

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```


### Tinorganics and PFAS
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <-  c("PFAS", "PFOS", "PFOSA",
             "DDEPP", "DDTEP", "DDTPP", "TDEPP")    # expected to NOT correlate with fat
             
df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

  
## Effect of fat, explained variance {.tabset}  
  
### PCBs
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c(
  "CB28", "CB52", "CB101", "CB105", 
  "CB118", "CB138", "CB153", "CB156", "CB180", "CB209", "CB_S7")

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE)) +
  geom_point(
    aes(y = Dev_expl_fat, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Deviance explained by fat (prop.)")

```

### BDEss
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c(
  "BDE28", "BDE47", "BDE49", 
  "BDE66", "BDE99", "BDE100", "BDE154", "BDESS", "BDE6S")

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE)) +
  geom_point(
    aes(y = Dev_expl_fat, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Deviance explained by fat (prop.)")

```


### Organochorines, paraffins
```{r, echo=FALSE, fig.width=6, fig.height=4}

params <- c("OCS", "HCB", "SCCP", "MCCP" )

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE)) +
  geom_point(
    aes(y = Dev_expl_fat, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Deviance explained by fat (prop.)")

```


### Tinorganics and PFAS
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <-  c("PFAS", "PFOS", "PFOSA",
             "DDEPP", "DDTEP", "DDTPP", "TDEPP")    # expected to NOT correlate with fat
df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE)) +
  geom_point(
    aes(y = Dev_expl_fat, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Deviance explained by fat (prop.)")

```

## Some more examples  

```{r, echo = FALSE}

get_data <- function(param, station, data){
  
  data %>%
    filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% param & 
             TISSUE_NAME %in% "Lever" & 
             STATION_CODE == station & 
             !is.na(FAT_PERC) & !is.na(LNMEA))
  
  }

```



```{r, echo = FALSE}
param <- "CB52"
station <- "36B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```



```{r, echo = FALSE}
param <- "CB118"
station <- "36B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "CB_S7"
station <- "36B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "BDE49"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "BDE99"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "BDE6S"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "OCS"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```



```{r, echo = FALSE}
param <- "HCB"
station <- "23B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```



```{r, echo = FALSE}
param <- "SCCP"
station <- "23B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```



```{r, echo = FALSE}
param <- "PFOS"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```



```{r, echo = FALSE}
param <- "DDEPP"
station <- "30B"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=9, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC) + LNMEA, data = df2)

par(mfrow = c(1,3), mar = c(4,5,2,1), oma = c(0,0,2,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0.5)

```

