---
title: "Trends in ww/dw concentration vs. trends in Fulton's condition index"
author: "DHJ"
date: "25 11 2020"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

*Fulton's index* (or *Fulton's condition factor* ) is given by K = 100(W/L<sup>3</sup>)   
See [Wikipedia](https://en.wikipedia.org/wiki/Standard_weight_in_fish#Applications)  

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)
library(plotly)
library(mgcv)
library(purrr)
library(visreg)
library(readxl)

```



### Number of time series
```{r, echo=FALSE}

df <- read_excel("Input_data/Longform assessment condition.xlsx", 
                 sheet = "DK SP NL", range = "A1:Z7862") %>%
  mutate(
    Conc_ww = case_when(
      Param_basis %in% "D" ~ `Param Conc`*DRYWT/100,
      Param_basis %in% "W" ~ `Param Conc`),
    Conc_dw = case_when(
      Param_basis %in% "D" ~ `Param Conc`,
      Param_basis %in% "W" ~ `Param Conc`/(DRYWT/100)),
    Parameter = factor(Parameter, 
                       levels = c("CD", "PB", "HG", "CB118", "CB153", "PA", "PYR"))
  )
  
# BOCHTVWTM

df %>%
  distinct(country, Parameter, station_name) %>%
  xtabs(~ country + Parameter, .)

```


```{r, fig.width=7, fig.height=9, echo=FALSE, results='hide'}

#
# Fixing some data - since we have no units
#

#
# Check
#

if (FALSE){
df %>%
  ggplot(aes(Conc_ww)) +
  geom_histogram(binwidth = 250) +
  facet_grid(col = vars(country), row = vars(Parameter))
}


### Histogram of ww values, PB in Spain  
# ```{r, fig.width=7, fig.height=5, echo=FALSE, results='hide'}

if (FALSE){
  
df %>%
  filter(Parameter == "PB" & country == "Spain") %>%
  ggplot(aes(Conc_ww)) +
  geom_histogram(binwidth = 100) +
  labs(title = "Original data")

df %>%
  filter(Parameter == "PB" & country == "Spain") %>%
  ggplot(aes(Conc_ww)) +
  geom_histogram() +
  scale_x_log10() +
  labs(title = "Original data, log scale")

}

#
# Put some Spanish data on same scale
#
df2 <- df %>%
  mutate(
    Conc_ww = case_when(
      Parameter == "PB" & country == "Spain" & Conc_ww > 1600 ~ Conc_ww/1000,
      TRUE ~ Conc_ww)
  )


#
# New histogram
#
if (FALSE){

  df2 %>%
  filter(Parameter == "PB" & country == "Spain") %>%
  ggplot(aes(Conc_ww)) +
  geom_histogram(binwidth = 100) +
  labs(title = "Fixed data")

}


```



## Example  

```{r, echo=FALSE}
station <- "BOCHTVWTM"
param <- "CD"
```
  

### `r param` at station `r station`  
```{r, echo = FALSE, warning=FALSE, fig.width = 5, fig.height=3}


df2 %>%
  filter(station_name == station & Parameter %in% param) %>%
  ggplot(aes(year, DRYWT)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, ", dry weight perentage by year"))

df2 %>%
  filter(station_name == station & Parameter %in% param) %>% 
  ggplot(aes(year, Fultons)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, ", Fulton's index by year"))

df2 %>%
  filter(station_name == station & Parameter %in% param) %>%
  ggplot(aes(year, Conc_ww)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, ", w.w. concentration by year"))


```


## Analyse trends        

```{r, echo = FALSE, warning=FALSE}

get_regressions <- function(param, station, data){
  
  # cat(param, station, "\n")
  df_select <- data %>%
    filter(station_name == station & Parameter %in% param & !is.na(Fultons) & !is.na(Conc_dw))
  
  modsumm1 <- summary(lm(Fultons ~ year, df_select))
  modsumm2 <- summary(lm(Conc_ww ~ year, df_select))
  modsumm3 <- summary(lm(Conc_dw ~ year, df_select))
  
  tibble(
    Parameter = param,
    station_name = station,
    country = df_select$country[1],
    fulton_slope = modsumm1$coef[2,1],
    fulton_se = modsumm1$coef[2,2],
    fulton_pval = modsumm1$coef[2,4],
    conc_ww_slope = modsumm2$coef[2,1],
    conc_ww_se = modsumm2$coef[2,2],
    conc_ww_pval = modsumm2$coef[2,4],
    conc_dw_slope = modsumm3$coef[2,1],
    conc_dw_se = modsumm3$coef[2,2],
    conc_dw_pval = modsumm3$coef[2,4]
  )
  
  
  
}

# Test
# get_regressions("CD", "BOCHTVWTM", df2)
 
```

```{r, echo = FALSE, warning = FALSE}

df_series <- df2 %>%
  filter(!is.na(Fultons) & !is.na(`Param Conc`)) %>%
  group_by(Parameter, station_name) %>%
  summarise(
    n_year = length(unique(year)), 
    .groups = "drop") %>%
  filter(n_year >= 5)   # at least 10 years and at least 70% data over LOQ

# xtabs(~station_name + Parameter, df_series)

df_slopes <- 1:nrow(df_series) %>%
  map_dfr(~get_regressions(df_series$Parameter[.],
                           df_series$station_name[.],
                           df2))


```

### W.w. conc. trend vs. Fulton trend   
* the x axis is the slope of the time trend of Fulton's index  
* the y axis is the slope of the time trend of w.w. concentration    
```{r, echo = FALSE, fig.width = 8, fig.height=6, warning=FALSE}

gg <- ggplot(df_slopes, aes(fulton_slope, conc_ww_slope, 
                            color = country, text = station_name)) +
  geom_point(size = 1) +
  facet_wrap(vars(Parameter), scales = "free_y")

ggplotly(gg)

```

### Ww trend minus dw trend, vs. Fulton trend   
* the x axis is the slope of the time trend of Fulton's index   
* the y axis is the *difference* in time trends (w.w. minus d.w.)      
```{r, echo = FALSE, fig.width = 8, fig.height=6, warning=FALSE}

gg <- df_slopes %>%
  mutate(ww_minus_dw_slope = conc_ww_slope - conc_dw_slope) %>%
  ggplot(aes(fulton_slope, ww_minus_dw_slope, 
             color = country, text = station_name)) +
  geom_point(size = 1) +
  facet_wrap(vars(Parameter), scales = "free_y")

ggplotly(gg)

```


### Ratio of Ww trend / dw trend, vs. Fulton trend    
* the x axis is the slope of the time trend of Fulton's index   
* the y axis is the *ratio* of time trends (w.w. divided by d.w.)      
```{r, echo = FALSE, fig.width = 8, fig.height=6}

gg <- df_slopes %>%
  mutate(ww_dw_slope_ratio = conc_ww_slope/conc_dw_slope) %>%
  ggplot(aes(fulton_slope, ww_dw_slope_ratio, 
             color = country, text = station_name)) +
  geom_point(size = 1) +
  facet_wrap(vars(Parameter), scales = "free_y")

ggplotly(gg)

```



## More examples   
  
In all plots,  
* DRYWT = dry weight percentage  
* Fultons = Fulton's index  
* Param Conc = Wet weight concentration
  

### Negative trend in both w.w. concentration and Fulton's index  

```{r, echo = FALSE}
param <- "CD"
station <- "MGACED"
```

`r param` at station `r station`  

```{r, echo = FALSE, warning=FALSE, fig.width = 5, fig.height=3}

df2_sel <- df2 %>%
  filter(station_name == station & Parameter %in% param)

ggplot(df2_sel, aes(year, DRYWT)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), dry weight perentage"))

ggplot(df2_sel, aes(year, Fultons)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), Fulton's index"))

ggplot(df2_sel, aes(year, `Param Conc`)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), w.w. concentration"))


```


### Negative trend in Fulton's index, no trend in w.w. concentration

```{r, echo = FALSE}
param <- "PB"
station <- "BOCHTVWTM"
```

`r param` at station `r station`  

```{r, echo = FALSE, warning=FALSE, fig.width = 5, fig.height=3}

df2_sel <- df2 %>%
  filter(station_name == station & Parameter %in% param)

ggplot(df2_sel, aes(year, DRYWT)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), dry weight perentage"))

ggplot(df2_sel, aes(year, Fultons)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), Fulton's index"))

ggplot(df2_sel, aes(year, `Param Conc`)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), w.w. concentration"))


```




### Negative trend in Fulton's index, positive trend in w.w. concentration

```{r, echo = FALSE}
param <- "CD"
station <- "NJY MSS11"
```

`r param` at station `r station`  

```{r, echo = FALSE, warning=FALSE, fig.width = 5, fig.height=3}

df2_sel <- df2 %>%
  filter(station_name == station & Parameter %in% param)

ggplot(df2_sel, aes(year, DRYWT)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), dry weight perentage"))

ggplot(df2_sel, aes(year, Fultons)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), Fulton's index"))

ggplot(df2_sel, aes(year, `Param Conc`)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), w.w. concentration"))


```

### Positive trend in Fulton's index, negative trend in w.w. concentration

```{r, echo = FALSE}
param <- "PB"
station <- "MGAPIN"
```

`r param` at station `r station`  

```{r, echo = FALSE, warning=FALSE, fig.width = 5, fig.height=3}

df2_sel <- df2 %>%
  filter(station_name == station & Parameter %in% param)

ggplot(df2_sel, aes(year, DRYWT)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), dry weight perentage"))

ggplot(df2_sel, aes(year, Fultons)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), Fulton's index"))

ggplot(df2_sel, aes(year, `Param Conc`)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  labs(title = paste0(param, " at ", station, " (", df2_sel$country[1], "), w.w. concentration"))


```

