---
title: "Correlations between fat and organics - Norwegian blue mussel"
author: "DHJ"
date: "25 11 2020"
output: 
  html_document:
    toc: true
    toc_float: true
---




```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

library(dplyr)
library(ggplot2)
# library(plotly)
library(mgcv)
library(purrr)
library(visreg)

```


```{r, echo = FALSE}

#
# Data (Norwegian CEMP data on sample level)
#

df <- readRDS("../Milkys2_pc/Files_from_Jupyterhub_2019/Raw_data/109_adjusted_data_2020-08-05.rds") %>%
  mutate(
    case_when(
    PARAM == "Cd" ~ "CD",
    TRUE ~ PARAM)
  )

# df %>%
#   filter(LATIN_NAME %in% "Gadus morhua" & TISSUE_NAME %in% "Lever") %>%
#   group_by(PARAM, STATION_CODE) %>%
#   summarise(n_year = length(unique(MYEAR)), over_loq = mean(is.na(FLAG1))) %>%
#   filter(n_year > 10 & over_loq > 0.7) %>% pull(PARAM) %>% unique() %>% dput()


```


```{r, echo = FALSE}
#
# Data series    
#
# Keep time series with at least 10 years and at least 70% data over LOQ
#
df_series <- df %>%
  filter(LATIN_NAME %in% "Mytilus edulis" & TISSUE_NAME %in% "Whole soft body" & !is.na(FAT_PERC)) %>%
  group_by(PARAM, STATION_CODE) %>%
  summarise(
    n_year = length(unique(MYEAR)), 
    over_loq = mean(is.na(FLAG1)),
    .groups = "drop") %>%
  filter(n_year > 15 & over_loq > 0.7)   # at least 10 years and at least 70% data over LOQ

# df_series$PARAM %>% unique() %>% dput()

sel_param <- c(
  "AG", "AS", "BA", "CD", "CO", "CR", "CU", "HG", "NI", "PB", 
  "CB_S7", "CB28", "CB52", "CB77", "CB81", "CB101", "CB105", 
  "CB118", "CB126", "CB138", "CB153", "CB156", "CB169", "CB180", "CB209",
  "HCB", "HCHG", "ICDP", 
  "TBT", "DBT", 
  "KPAH", "P_S", "PAH16", "ACNE", "ANT", "BAP", "BBJF", "BEP", "BGHIP", 
  "BKF", "BAA",  "CHR", "FLE", "FLU", "PA", "PER", "PYR",
  "DDEPP", "DDTEP", "DDTPP", "TCDD", "TDEPP"
)

df_series <- df_series %>%
  filter(PARAM %in% sel_param) %>%
  mutate(PARAM = factor(PARAM, levels = sel_param)) %>%
  arrange(PARAM, STATION_CODE)

```

```{r, echo = FALSE}

param <- "CB52"
station <- "52A"

```


## Example: one substance (`r param`), one station (`r station`)    
  
Model: `ln(concentration) = a*ln(fat) + spline(Year)`  

Thus:  

* The effect of ln(fat) assumed to be linear. This effect is given by the line `log(FAT_PERC)` in the output below. `Estimate` gives the slope of the effect      
* The effect of year (non-linear effect) is also taken into account, i.e. it is controlled for general cganges in the concentration (given by the line `s(MYEAR)` in the output)  
```{r, echo = FALSE, fig.height=4, fig.width=8}

# param <- "CB_S7"


#
# Test 1 (result not shown): one substance, all stations    
#
df2 <- df %>%
  filter(LATIN_NAME %in% "Mytilus edulis" & PARAM %in% param & 
           TISSUE_NAME %in% "Whole soft body" & !is.na(FAT_PERC))


#
# Test 2 (shown): one substance + one station
#

df3_check <- df2 %>%
  filter(STATION_CODE == station)

# All stations    
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df3_check)
summary(m1)

par(mfrow = c(1,2), mar = c(4,5,1,1), oma = c(0,0,1.5,0))
visreg(m1)
mtext(paste0(param, ", ", station), outer = TRUE, line = 0)

# summary(m1) %>% str()
# summary(m1)$p.table[2,]

```


### Several substances + stations          
```{r, echo = FALSE}

#
# Test with one substance + one station
#

# Test with one substance  

get_estimates <- function(param, station, data){
  
  # For debugging:
  # cat(param, station, "\n")
  
  df2 <- data %>%
    filter(LATIN_NAME %in% "Mytilus edulis" & PARAM %in% param & 
           TISSUE_NAME %in% "Whole soft body" & !is.na(FAT_PERC) &
             STATION_CODE == station & PARAM %in% param)

  m0 <- gam(log(VALUE_WW) ~ 1, data = df2)
  m1 <- gam(log(VALUE_WW) ~ s(MYEAR), data = df2)
  m2 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)
  dev <- anova(m0, m1, m2)[,2]
  
  tibble(PARAM = param, 
         STATION_CODE = station,
         Fat_slope = summary(m2)$p.table[2,1],
         Fat_slope_se = summary(m2)$p.table[2,2],
         Fat_pvalue = summary(m2)$p.table[2,4],
         Dev_expl_tot = (dev[1]-dev[3])/dev[1],
         Dev_expl_fat = (dev[2]-dev[3])/dev[1]
         )
  
}

# Test
# debugonce(get_estimates)
# get_estimates("CB52", "52A", df)
# get_estimates("AS", "B2", df)


```

```{r, echo = FALSE}
#
# Run analysis for each time series
#
df_effect <- 1:nrow(df_series) %>% 
  purrr::map_dfr(
    ~get_estimates(param = df_series$PARAM[.], station = df_series$STATION_CODE[.], df)
  )

df_effect <- df_effect %>%
  mutate(Fat_effect = ifelse(Fat_pvalue < 0.05, "P < 0.05",  "P >= 0.05"))

```


## Effect of fat, slope {.tabset}  

Effect of ln(fat) on ln(concentration)   

* Analysis on station-by-station basis   
* The effect of ln(fat) assumed to be linear   
* Effect of year (non-linear effct) and length (linear) also taken into account   
* Analysis done for time series with at least 10 years and at least 70% data over LOQ    
  
### PCBs
```{r, echo=FALSE, fig.width=10, fig.height=9}

params <- c(
  "CB28", "CB52", "CB101", "CB105", 
  "CB118", "CB138", "CB153", "CB156", "CB180", "CB209", "CB_S7")

df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM), ncol = 3) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, size = 6)) +
  labs(y = "Effect of fat on ln ww-concentration (slope)")

```






### DDT, TBT and others
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <-  c("DDEPP", "DDTEP", "DDTPP", "TDEPP",
             "HCB", "HCHG", "ICDP", "TBT", "DBT")    # expected to NOT correlate with fat
             
df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, size = 6)) + 
  labs(y = "Effect of fat on ln ww-concentration (slope)")


```

### PAHs
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c("KPAH", "P_S", "PAH16", "ACNE", "ANT", "BAP", "BBJF", "BEP", "BGHIP", 
  "BKF", "BAA",  "CHR", "FLE", "FLU", "PA", "PER", "PYR")
             
df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, size = 9)) + 
  labs(y = "Effect of fat on ln ww-concentration (slope)")


```

### Metals  
```{r, echo=FALSE, fig.width=9, fig.height=6}

params <- c(  "AG", "AS", "BA", "CD", "CO", "CR", "CU", "HG", "NI", "PB")
             
df_effect %>%
  filter(PARAM %in% params) %>%
  mutate(PARAM = factor(PARAM, levels = params)) %>%
  ggplot(aes(x = STATION_CODE, y = Fat_slope)) +
  geom_pointrange(
    aes(ymin = Fat_slope - 2*Fat_slope_se, ymax = Fat_slope + 2*Fat_slope_se, color = Fat_effect)) +
  scale_color_manual(values = c("red3", "grey40")) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PARAM)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, size = 6)) +
  labs(y = "Effect of fat on ln ww-concentration (slope)")


```


## Some more examples  

```{r, echo = FALSE}

get_data <- function(param, station, data){
  
  data %>%
    filter(LATIN_NAME %in% "Mytilus edulis" & PARAM %in% param & 
             TISSUE_NAME %in% "Whole soft body" & 
             STATION_CODE == station & 
             !is.na(FAT_PERC) )
  
  }

```



```{r, echo = FALSE}
param <- "CB52"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "CB118"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "DDEPP"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "TBT"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```


```{r, echo = FALSE}
param <- "PYR"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```




```{r, echo = FALSE}
param <- "PYR"
station <- "I131A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```




```{r, echo = FALSE}
param <- "CU"
station <- "30A"
```

### `r param` at station `r station`      
```{r, echo = FALSE, fig.width=6, fig.height=3}

df2 <- get_data(param, station, df)
m1 <- gam(log(VALUE_WW) ~ s(MYEAR) + log(FAT_PERC), data = df2)

par(mfrow = c(1,2), mar = c(4,5,0,1), oma = c(0,0,2,0))
visreg(m1, xlab = "Year", ylab = "ln ww. concentration")
mtext(paste0(param, ", station ", station), outer = TRUE, line = 0.5)

```

